---
title: "Arabidopsis_enrich_gene_visualization"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---

# Introduction #
This is a Notebook for enrichment ratio results visualization. 
The output from a [custom python script](https://github.com/guoweier/Poplar_Chromoanagenesis/blob/master/enrichment-ratio/enrichment-window-ratio.py) shows: 
1. Target genomic feature density surround real breakpoints;
2. Target genomic feature density surround 10,000 (default number) pseudo breakpoints, with 1,000 (default number) random selective pseudo population;
3. One t-test results displaying the p-value with each pseudo population and the real breakpoints group.

This R script is aimed to visualize these data into density plot, by using ggplot package. 

# Outline #
1. Input dataset and libraries
2. 

# Input libraries and dataset #
## input library ##
```{r}
library(ggplot2)
library("ggpubr")
```

## input 1kb window dataset ##
### input psuedo dataset ###
```{r}
DHS.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_DHS_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
gene.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_genes_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
mRNA.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_mRNA_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
pseudo.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_pseudogene_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
ori.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_replication_origins_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT1.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state1_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT2.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state2_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT3.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state3_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT4.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state4_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT5.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state5_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT6.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state6_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT7.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state7_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT8.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state8_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT9.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_state9_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
TE.rand.1k <- read.table("../enrich/input/Gaetan_aa_C_transposable_element_1k_20220614-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
```

### adjust parameters in pseudo dataset ###
```{r}
mean.1k = ori.rand.1k$Random.mean
pvalue.1k = ori.rand.1k$P.value
mu.1k = mean(mean.1k)
avg_pvalue.1k = mean(pvalue.1k)
df.rand.1k = data.frame(ori.rand.1k)
head(df.rand.1k)
```

### input breakpoints dataset ###
```{r}
DHS.real.1k = read.table("../enrich/input/Arabidopsis-DHS-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
gene.real.1k = read.table("../enrich/input/Arabidopsis-genes-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
mRNA.real.1k = read.table("../enrich/input/Arabidopsis-mRNA-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
pseudo.real.1k = read.table("../enrich/input/Arabidopsis-pseudogene-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
ori.real.1k = read.table("../enrich/input/Arabidopsis-replication_origins-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT1.real.1k = read.table("../enrich/input/Arabidopsis-state1-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT2.real.1k = read.table("../enrich/input/Arabidopsis-state2-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT3.real.1k = read.table("../enrich/input/Arabidopsis-state3-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT4.real.1k = read.table("../enrich/input/Arabidopsis-state4-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT5.real.1k = read.table("../enrich/input/Arabidopsis-state5-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT6.real.1k = read.table("../enrich/input/Arabidopsis-state6-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT7.real.1k = read.table("../enrich/input/Arabidopsis-state7-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT8.real.1k = read.table("../enrich/input/Arabidopsis-state8-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT9.real.1k = read.table("../enrich/input/Arabidopsis-state9-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
TE.real.1k = read.table("../enrich/input/Arabidopsis-transposable_element-1kw-1kd-20220614.txt", sep="\t", header = TRUE, stringsAsFactors = F)
```

### adjust parameters in real breakpoints dataset ###
```{r}
enrich.1k = ori.real.1k$EnrichmentRatio
realmean.1k = mean(enrich.1k)
realmin.1k = min(enrich.1k)
realmax.1k = max(enrich.1k)
```

## compare pseudo dataset and real breakpoints dataset ##
```{r}
enratio.1k = realmean.1k/mu.1k
enratio.1k
avg_pvalue.1k
realmean.1k
```

# histogram with density plot with mean line #
```{r}
meanline.1k <- ggplot(df.rand.1k, aes(x=mean.1k)) +
  geom_histogram(colour="black", fill="black", bins = 120) +
  geom_vline(aes(xintercept=realmean.1k, color="blue"), size=2, show.legend = FALSE) +
  scale_x_continuous(name = "Mean replication origin density") +
  scale_y_continuous(name = "Number of Breakpoints") +
  labs(title = "Replication origin density in 1kb window") +
  theme(axis.text = element_text(size = 10), axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12), plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
meanline.1k
```

# draw text for p-value and ratio info #
```{r}
text1 <- paste("Enrichment Ratio = 1.50\nP-value = 0.047*")
text1.p <- ggparagraph(text = text1, color = "black", size = 8)
```

# add statistical info to plot #
```{r}
den.text.1k <- meanline.1k + annotation_custom(ggplotGrob(text1.p),xmin = 0.05, xmax = 0.2, ymax = 45)
den.text.1k
```


# save 1k window plot #
```{r}
ggsave("../enrich/output/ORI-1kwindow-20220616.png", width = 5, height = 3)
```



## input 10kb window dataset ##
### input psuedo dataset ###
```{r}
DHS.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_DHS_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
gene.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_genes_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
mRNA.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_mRNA_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
pseudo.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_pseudogene_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
ori.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_replication_origins_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT1.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state1_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT2.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state2_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT3.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state3_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT4.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state4_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT5.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state5_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT6.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state6_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT7.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state7_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT8.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state8_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
CT9.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_state9_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
TE.rand.10k <- read.table("../enrich/input/Gaetan_aa_C_transposable_element_10k_20220616-Stats.txt",sep="\t", header = TRUE, stringsAsFactors = F)
```

### adjust parameters in pseudo dataset ###
```{r}
mean.10k = TE.rand.10k$Random.mean
pvalue.10k = TE.rand.10k$P.value
mu.10k = mean(mean.10k)
avg_pvalue.10k = mean(pvalue.10k)
df.rand.10k = data.frame(TE.rand.10k)
head(df.rand.10k)
```

### input breakpoints dataset ###
```{r}
DHS.real.10k = read.table("../enrich/input/Arabidopsis-DHS-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
gene.real.10k = read.table("../enrich/input/Arabidopsis-genes-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
mRNA.real.10k = read.table("../enrich/input/Arabidopsis-mRNA-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
pseudo.real.10k = read.table("../enrich/input/Arabidopsis-pseudogene-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
ori.real.10k = read.table("../enrich/input/Arabidopsis-replication_origins-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT1.real.10k = read.table("../enrich/input/Arabidopsis-state1-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT2.real.10k = read.table("../enrich/input/Arabidopsis-state2-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT3.real.10k = read.table("../enrich/input/Arabidopsis-state3-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT4.real.10k = read.table("../enrich/input/Arabidopsis-state4-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT5.real.10k = read.table("../enrich/input/Arabidopsis-state5-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT6.real.10k = read.table("../enrich/input/Arabidopsis-state6-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT7.real.10k = read.table("../enrich/input/Arabidopsis-state7-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT8.real.10k = read.table("../enrich/input/Arabidopsis-state8-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
CT9.real.10k = read.table("../enrich/input/Arabidopsis-state9-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
TE.real.10k = read.table("../enrich/input/Arabidopsis-transposable_element-10kw-1kd-20220616.txt", sep="\t", header = TRUE, stringsAsFactors = F)
```

### adjust parameters in real breakpoints dataset ###
```{r}
enrich.10k = TE.real.10k$EnrichmentRatio
realmean.10k = mean(enrich.10k)
realmin.10k = min(enrich.10k)
realmax.10k = max(enrich.10k)
```

## compare pseudo dataset and real breakpoints dataset ##
```{r}
enratio.10k = realmean.10k/mu.10k
enratio.10k
avg_pvalue.10k
realmean.10k
```

# histogram with density plot with mean line #
```{r}
meanline.10k <- ggplot(df.rand.10k, aes(x=mean.10k)) +
  geom_histogram(colour="black", fill="black", bins = 120) +
  geom_vline(aes(xintercept=realmean.10k, color="blue"), size=2, show.legend = FALSE) +
  scale_x_continuous(name = "Mean transposable element density") +
  scale_y_continuous(name = "Number of Breakpoints") +
  labs(title = "Transposable element density in 10kb window") +
  theme(axis.text = element_text(size = 10), axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12), plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
meanline.10k
```

# draw text for p-value and ratio info #
```{r}
text1 <- paste("Enrichment Ratio = 0.42\nP-value = 2.44E-53***")
text1.p <- ggparagraph(text = text1, color = "black", size = 8)
```

# add statistical info to plot #
```{r}
den.text.1k <- meanline.10k + annotation_custom(ggplotGrob(text1.p),xmin = 0.12, xmax = 0.2, ymax = 125)
den.text.1k
```

# save 10k window plot #
```{r}
ggsave("../enrich/output/TE-10kwindow-20220616.png", width = 5, height = 3)
```
