---
title: "Arabidopsis_SNP_freq_visualization"
output:
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
---

# Introduction #
The Notebook is aimed to visualize the SNP frequency (Col vs Ler) in asy1 shattered hybrids.
The SNP% is for Ler. The SNP% would be different specifically around centromeres coming from different non-disjunctions.

- If shattered chromosome from non-disjunction in Meiosis I:
  with shattered arm: 33%; no shattered arm: 0% or 50%
- if shattered chromosome from non-disjunction in Meiosis II:
  with shattered arm: 66% or 0%; no shattered arm: 0% or 50%

# Outline #
1. load libraries and data
2. 

# Input libraries and data #
## add libraries ##
```{r}
library("tidyverse")
library("ggplot2")
library("ggpubr")
```


## add data ##
### 10kb ###
```{r}
asy1.SNP.10kb <- read.csv("/Users/wendy/Desktop/Lab/arabidopsis/binbygeno_asy1_10kb_20220310.txt",sep = "\t")
head(asy1.SNP.10kb)
```


### 100kb ###
```{r}
asy1.SNP.100kb <- read.csv("/Users/wendy/Desktop/Lab/arabidopsis/binbygeno_asy1_100kb_20220310.txt",sep = "\t")
head(asy1.SNP.100kb)
```

# take data in Chr1 #
## 10kb ##
```{r}
pos.chr1.10kb <- asy1.SNP.10kb$Start[asy1.SNP.10kb$Chrom == 'Chr1']
ler.chr1.10kb <- asy1.SNP.10kb$v.Gaetan.aa.C.102.Obs.B[asy1.SNP.10kb$Chrom == 'Chr1']
chr1.10kb <- data.frame(pos.chr1.10kb,ler.chr1.10kb)
snp.chr1.10kb <- list(pos.chr1.10kb)
```

# Chr1 #
## 100kb ##
```{r}
pos.chr1.100kb <- asy1.SNP.100kb$Start[asy1.SNP.100kb$Chrom == 'Chr1']
ler.chr1.100kb <- asy1.SNP.100kb$v.Gaetan.aa.C.102.Obs.B[asy1.SNP.100kb$Chrom == 'Chr1']
chr1.100kb <- data.frame(pos.chr1.100kb, ler.chr1.100kb)
```


# SNP frequency plots #
## 10kb ##
```{r}
snp.10kb <- ggplot(chr1.10kb, aes(pos.chr1.10kb,ler.chr1.10kb)) +
  geom_point(alpha=.6) +
  scale_color_manual(values = c("#E69F00")) +
  theme(axis.text.x = element_text(size = 8, angle = 90), axis.title.x = element_text(size = 14, margin = margin(t=10)), axis.title.y = element_text(size = 14, margin = margin(r=10)), legend.position = "None") + 
  scale_x_continuous(name="Chromosome 1", breaks = NULL) +
  scale_y_continuous(name="Ler SNP Frequency")
snp.10kb
```


# save #
## 10kb ##
```{r}
ggsave("/Users/wendy/Desktop/Lab/arabidopsis/asy1_snp_chr1_10kb_20220310.png",width = 10, height = 4)
```

# dosage data #
## 5kb only chr1 ##
```{r}
asy1.dos.10k <- read.csv("/Users/wendy/Desktop/Lab/arabidopsis/binbysam-Gaetan-102-2ctrl-10kb-20210916.txt",sep = "\t")
asy1.strt <- asy1.dos.10k$Strt[asy1.dos.10k$Chrom == 'Chr1']
asy1.Cov <- asy1.dos.10k$Gaetan_aa_C_102.LC_C_1_21_ctrl[asy1.dos.10k$Chrom == 'Chr1']
asy1.dos.Chr1 <- data.frame(asy1.strt, asy1.Cov)
head(asy1.dos.Chr1)
```
## compare dos and snp locations ##
```{r}
dos.chr1.strt.new <- c()
dos.chr1.cov.new <- c()
for (a in asy1.dos.Chr1$asy1.strt) {
  pos <- a-1
  for (b in snp.chr1.10kb[[1]]) {
    if (pos == b) {
      dos.chr1.strt.new <- c(dos.chr1.strt.new, pos)
      dos.chr1.cov.new <- c(dos.chr1.cov.new, asy1.dos.Chr1$asy1.Cov[asy1.strt == a])
    }
  }
}
length(dos.chr1.strt.new)
length(dos.chr1.cov.new)
length(snp.chr1.10kb[[1]])
```

## combine snp and dos with same positions ##
```{r}
dos.chr1.cov.new <- as.numeric(dos.chr1.cov.new)
snp.dos.chr1.10k <- data.frame(dos.chr1.strt.new, dos.chr1.cov.new, ler.chr1.10kb)
head(snp.dos.chr1.10k)
```

# save file #
```{r}
write.table(snp.dos.chr1.10k, file = "/Users/wendy/Desktop/Lab/arabidopsis/snp_dos_chr1_10kb_20220315.txt", sep = "\t", row.names = TRUE, col.names = TRUE)
```


# plot dos and snp together #
```{r}
dos.snp.10k <- ggplot(snp.dos.chr1.10k, aes(x = dos.chr1.strt.new)) +
  geom_point(aes(y=dos.chr1.cov.new),alpha=.7,color="black") +
  geom_point(aes(y=ler.chr1.10kb/25),alpha=.5,color="navy") +
  theme(axis.text.x = element_text(size = 8, angle = 90), axis.title.x = element_text(size = 14, margin = margin(t=10)), axis.title.y = element_text(size = 14, margin = margin(r=10))) +
  scale_x_continuous(name = "Chromosome 1", breaks = NULL) +
  scale_y_continuous(name = "Relative Read Coverage",sec.axis = sec_axis(trans = ~.*25, name = "Ler SNP Frequency"))
dos.snp.10k
```

# save plot #
```{r}
ggsave("/Users/wendy/Desktop/Lab/arabidopsis/snp_dos_chr1_10kb_20220315.png", height = 4, width = 8)
```


# dosage plot #
```{r}
dos.10kb <- ggplot(asy1.dos.Chr1, aes(asy1.strt,as.numeric(asy1.Cov))) +
  geom_point(alpha=.6) +
  theme(axis.text.x = element_text(size = 8, angle = 90), axis.title.x = element_text(size = 14, margin = margin(t=10)), axis.title.y = element_text(size = 14, margin = margin(r=10)), legend.position = "None") +
  scale_x_continuous(name = "Chromosome 1", breaks = NULL) +
  scale_y_continuous(name = "Relative Read \nCoverage",limits = c(1,4))
dos.10kb
```

# arrange two plots into one #
```{r}
onepage <- ggarrange(snp.10kb,dos.10kb,labels = c("A","B"), ncol = 1, nrow = 2)
onepage
```

# save onepage #
```{r}
ggsave("/Users/wendy/Desktop/Lab/arabidopsis/asy1_snp_dos_10kb_20220310.png", width = 5.5, height = 5)
```




